Radio Crickets: Chirping Jets from Black Hole Binaries Entering their Gravitational Wave Inspiral

### 2015-07-07 Paper/paper.tex
Update abstract

\begin{abstract}
Apart from a stochastic gravitational wave (GW) background, eLISA will
also be able to detect GW emission from binary black holes.  Combined
GW and electromagnetic observations of compact binary mergers should
enable detailed studies of astrophysical processes in the strong-field
gravity regime.  Here we show that if one of component BHs of the
binary has an accretion disk with an associated jet, the morphology of
the jet can act as an electromagnetic counterpart to the GW signal
from the binary.  For binaries with mass $10^7$--$10^{10}$ M$_\odot$
at redshifts $z<0.5$, the jet precesses along a biconical surface due
to the binary's orbital motion.  When the binary enters the GW phase
of its evolution, the opening angle widens, the jet exhibits
milliarcsecond scale wiggles and the conical surface of jet precession
is twisted due to the superluminal motion.  These morphological
features can be used to constrain the dyanamical parameters of the
binary.  Monitoring of jets with these features in current and
archival data will put a lower limit on the sources that could be
detected by GW experiments such as eLISA and Pulsar Timing Arrays
(PTAs).  In the future, microarcsecond interferometry with experiments
such as SKA will increase the potential usefulness of this technique.  
\end{abstract}

>>>>

\begin{abstract}
We study a novel electromagnetic signature of supermassive black hole
binaries whose inspiral stars being dominated by gravitational wave
(GW) emission. Combined GW and electromagnetic observations of compact
binary mergers would allow to use them as standard sirens for
cosmology and a testing ground of general relativity.  If one of the
binary members produces a radio jet, the helical chirping morphology
of the jet can be used to infer the binary parameters.  For binaries
with mass $10^7$--$10^{10}$ M$_\odot$ at redshifts $z<0.5$, the jet
precesses along a biconical surface due to the binary's orbital
motion.  When the binary enters the GW phase of its evolution, the
opening angle widens, the jet exhibits milliarcsecond scale wiggles
and the conical surface of jet precession is twisted due to the
superluminal motion.  Monitoring these features in current and
archival data will place a lower limit on sources that could be
detected by eLISA and Pulsar Timing Arrays.  In the future,
microarcsecond interferometry with the Square Kilometer Array will
increase the potential usefulness of this technique.
\end{abstract}
